---
title: "viewing"
author: "Keiryn Hart, 300428418"
date: "21/08/2020"
output: html_document
---

# reading the datasets into r, separating the victims dataset by day,month and year again so I can subset the new victims dataset by specific months. from this I have given the victims dataset theoretical alert levels for comparison purposes with the covid dataset.
```{r}
library (plyr)
library(tidyr)

victims <- read.csv("/Users/keirynhart/Documents/Uni/Data 301/DATA301-Group-Project/Data/Victimisations_time_and_place (cleaned).csv", header = TRUE)
covid <- read.csv("/Users/keirynhart/Documents/Uni/Data 301/DATA301-Group-Project/Data/Covid calls for service cleaned.csv", header = TRUE)

victims <- separate(victims, "date", c("day", "month", "year"))           

months <- c(3,4,5)
victims <- subset(victims, month %in% months)

victims[, "theoretical_alert_level"] <- NA

for(i in 1:nrow(victims)){
  if(victims$month[i] == 3){
    victims$theoretical_alert_level[i] <- 4
  }
  else if(victims$month[i] == 4){
    victims$theoretical_alert_level[i] <- 3
  }
  else if(victims$month[i] == 5){
    victims$theoretical_alert_level[i] <- 2
  }
}
```

#Here I am averaging the Locations from the covid dataset into regions that can be matched to the victims dataset for merging.
```{r}
covid[, "Location"] <- NA

for(i in 1:nrow(covid)){
  if(covid$District[i] == "Auckland City" | covid$District[i] == "Counties/Manukau" | covid$District[i] == "Waitemata"){
    covid$Location[i] <- "Auckland"
  }
  else if(covid$District[i] == "Bay Of Plenty"){
    covid$Location[i] <- "Bay Of Plenty"
  }
  else if(covid$District[i] == "Waikato"){
    covid$Location[i] <- "Waikato"
  }
  else if(covid$District[i] == "Eastern"){
    covid$Location[i] <- "Hawkes Bay/ Gisborne"
  }
  else if(covid$District[i] == "Central"){
    covid$Location[i] <- "West Coast/ Upper-Central North"
  }
  else if(covid$District[i] == "Wellington"){
    covid$Location[i] <- "Wellington/ Lower-Central North"
  }
  else if(covid$District[i] == "Tasman"){
    covid$Location[i] <- "Tasman/ South-West Coast"
  }
  else if(covid$District[i] == "Northland"){
    covid$Location[i] <- "Northland"
  }
  else if(covid$District[i] == "Southern"){
    covid$Location[i] <- "Southland/ Otago"
  }
  else if(covid$District[i] == "Canterbury"){
    covid$Location[i] <- "Cantebury"
  }
  else{covid$Location[i] <- "Other"}
}

head(covid)
```

Again averaging the Locations in the victims dataset now to have a consistent match between the two for comparison.
```{r}
victims[, "Location"] <- NA

for(i in 1:nrow(victims)){
  if(victims$Territorial.Authority[i] == "Auckland."){
    victims$Location[i] <- "Auckland"
  }
  else if(victims$Territorial.Authority[i] == "Far North District." | victims$Territorial.Authority[i] == "Whangarei District." | victims$Territorial.Authority[i] == "Kaipara District."){
    victims$Location[i] <- "Northland"
  }
  else if(victims$Territorial.Authority[i] == "Waipa District." | victims$Territorial.Authority[i] == "Hamilton City." | victims$Territorial.Authority[i] == "Taupo District." | victims$Territorial.Authority[i] == "Hauraki District." | victims$Territorial.Authority[i] == "Otorohanga District." | victims$Territorial.Authority[i] == "South Waikato District." | victims$Territorial.Authority[i] == "Thames-Coromandel District." | victims$Territorial.Authority[i] == "Waitomo District." | victims$Territorial.Authority[i] == "Waikato District." | victims$Territorial.Authority[i] == "Matamata-Piako District."){
    victims$Location[i] <- "Waikato"
  }
  else if(victims$Territorial.Authority[i] == "Whakatane District." | victims$Territorial.Authority[i] == "Opotiki District." | victims$Territorial.Authority[i] == "Kawerau District." | victims$Territorial.Authority[i] == "Tauranga City." | victims$Territorial.Authority[i] == "Rotorua District." | victims$Territorial.Authority[i] == "Western Bay of Plenty District."){
    victims$Location[i] <- "Bay Of Plenty"
  }
  else if(victims$Territorial.Authority[i] == "Napier City." | victims$Territorial.Authority[i] == "Central Hawke's Bay District." | victims$Territorial.Authority[i] == "Hastings District." | victims$Territorial.Authority[i] == "Gisborne District." | victims$Territorial.Authority[i] == "Wairoa District."){
    victims$Location[i] <- "Hawkes Bay/ Gisborne"
  }
  else if(victims$Territorial.Authority[i] == "South Taranaki District." | victims$Territorial.Authority[i] == "Stratford District." | victims$Territorial.Authority[i] == "Ruapehu District." | victims$Territorial.Authority[i] == "Rangitikei District." | victims$Territorial.Authority[i] == "New Plymouth District." | victims$Territorial.Authority[i] == "Manawatu District." | victims$Territorial.Authority[i] == "Wanganui District." | victims$Territorial.Authority[i] == "Palmerston North City."){
    victims$Location[i] <- "West Coast/ Upper-Central North"
  }
  else if(victims$Territorial.Authority[i] == "Masterton District." | victims$Territorial.Authority[i] == "South Wairarapa District." | victims$Territorial.Authority[i] == "Lower Hutt City." | victims$Territorial.Authority[i] == "Porirua City." | victims$Territorial.Authority[i] == "Kapiti Coast District." | victims$Territorial.Authority[i] == "Horowhenua District." | victims$Territorial.Authority[i] == "Wellington City."  | victims$Territorial.Authority[i] == "Tararua District." | victims$Territorial.Authority[i] == "Carterton District." | victims$Territorial.Authority[i] == "Upper Hutt City."){
    victims$Location[i] <- "Wellington/ Lower-Central North"
  }
  else if(victims$Territorial.Authority[i] == "Grey District." | victims$Territorial.Authority[i] == "Buller District." | victims$Territorial.Authority[i] == "Tasman District." | victims$Territorial.Authority[i] == "Marlborough District." | victims$Territorial.Authority[i] == "Westland District." | victims$Territorial.Authority[i] == "Nelson City."){
    victims$Location[i] <- "Tasman/ South-West Coast"
  }
  else if(victims$Territorial.Authority[i] == "Invercargill City." | victims$Territorial.Authority[i] == "Queenstown-Lakes District." | victims$Territorial.Authority[i] == "Southland District." | victims$Territorial.Authority[i] == "Central Otago District." | victims$Territorial.Authority[i] == "Gore District." | victims$Territorial.Authority[i] == "Clutha District." | victims$Territorial.Authority[i] == "Dunedin City."){
    victims$Location[i] <- "Southland/ Otago"
  }
  else{victims$Location[i] <- "Canterbury"}
}
```


# removing unneccessary columns and preparing for merge.
```{r}
victims[, c("X", "X.1", "day")] <- NULL 
covid[,c("X", "Calls.for.Service.and.Prevention", "Crime.and.NonCrime.Demand", "Event.Category", "District")] <- NULL

covid[, "year"] <- 20

library(dplyr)
library(plyr)
covid <- covid %>% dplyr::rename(Event = Event.Type)

victims <- victims %>% dplyr::rename(Event = ANZSOC.Group, Alert.Level = theoretical_alert_level)

colnames(covid)
colnames(victims)

victims[,c("Territorial.Authority", "Area.Unit", "month")] <- NULL

covid[, "Meshblock"] <- NULL
victims[, "number_of_victimizations"] <- NA

victims[, c("Meshblock", "number_of_victimizations")] <- NULL

```

#creating matching sums for total events in both datasets for the merge.
```{r}
VictimsSum <- dplyr::count(victims, Event, year, Alert.Level, Location) %>% ungroup()

CovidSum <- aggregate(. ~  Event + year + Alert.Level + Location, covid, sum)

VictimsSum <- rename(VictimsSum, c("n" = "Number of Victimizations"))
CovidSum <- rename(CovidSum, c("No" = "Number of Victimizations"))
```

merge.
```{r}
total <- rbind(VictimsSum, CovidSum)

write.csv(total,"/Users/keirynhart/Documents/Uni/Data 301/DATA301-Group-Project/Data/merged data (Final).csv")
```
